---
title: Introduction to Social Network Analysis with igraph
author: "Robert Ackland & Timothy Graham"
date: "23 June 2017"
output: pdf_document
graphics: yes
urlcolor: blue
---

##Introduction

This tutorial provides an introduction to network analysis, using networks collected via the *SocialMediaLab* R package (for more information on *SocialMediaLab*, see the [page on the VOSON website](http://vosonlab.net/SocialMediaLab)).

The installation of *SocialMediaLab* and collection of data from Twitter, Facebook and YouTube is covered in an earlier tutorial.

Network analysis is conducted using the [igraph](https://cran.r-project.org/web/packages/igraph/index.html) R package, maintained by Gabor Csardi.


In this session we conduct basic social network analysis using *igraph*

##Setting up

First load the *igraph* package

```{r}
library(igraph)
```

Now read in the graphml file.  This YouTube comment network was collected using *SociaMediaLab* using the following 5 videos on "Fracking Australia":

https://www.youtube.com/watch?v=XrE7LzZCn1E

https://www.youtube.com/watch?v=PELxZ3K2o0c

https://www.youtube.com/watch?v=Xfo0hpVrtrs

https://www.youtube.com/watch?v=7Toi2s05Exg

https://www.youtube.com/watch?v=BBSVLGf7zPI

```{r}
g <- read.graph("http://vosonlab.net/papers/Taiwan_2017/Sat_Jun_27_13-04-34_2015_EST_YoutubeActorNetwork_FiveFracking.graphml", format="graphml")
#g <- read.graph("Sat_Jun_27_13-04-34_2015_EST_YoutubeActorNetwork_FiveFracking.graphml", format="graphml")
```

##Some basic information on the network

```{r}
g   #Information on the igraph graph object
```

### bipartite graph 可以分成兩群點，兩群點之間有邊，兩群點內部無邊
This informs us that there are 137 nodes and 359 edges in the network. It tells us that our graph is *D*irected, *N*amed, the edges are *W*eighted, and it is not a *B*ipartite graph.

```{r}
V(g)               #list of vertices
E(g)               #list of edges
V(g)[36]           #accessing particular vertex
E(g)[1]            #accessing particular edge

V(g)$id            #list of "id" (vertex) attributes
E(g)$weight        #list of "weight" (edge) attributes

vcount(g)          #number of nodes in network
length(V(g))       #another way

ecount(g)          #number of edges
length(E(g))       #another way

list.vertex.attributes(g)    #list of the vertex attributes
list.edge.attributes(g)      #list of the edge attributes
```

##Network maps

```{r warning=FALSE}
#png("csg_plot1.png", width=800, height=700)
plot(g, vertex.shape="none",edge.width=1.5, edge.curved = .5, edge.arrow.size=0.5, asp=9/16,margin=-0.15)
#dev.off()
```

### 圖中有 loop (Graph Theory 中，loop 是單點自己連到自己，cycle 是多點形成一個循環)
### 這是因為你有可能回應自己的 post 所導致，所以我們要把這種狀況拿掉
We notice that there are loops in the network, let's investigate this further...

```{r warning=FALSE}
is.simple(g)                 #test whether "simple" graph (no loops or multiple edges)

g2 <- simplify(g)            #construct "simple" version of network
g2
is.simple(g2)

#png("csg_plot2.png", width=800, height=700)
plot(g2, vertex.shape="none",edge.width=1.5, edge.curved = .5, edge.arrow.size=0.5, asp=9/16,margin=-0.15)
#dev.off()
```

### 查看連結性 (強連結或弱連結)
##Looking at the connectivity of the graph

```{r warning=FALSE}
neighbors(g2,36)                     #Leo Sayer's network neighbors

is.connected(g2, mode="weak")        #this is not a weakly connected component 

cc <- clusters(g2)             #information on connected components
head(cc$membership)                  #which component each node is assigned to
cc$csize                       #size of each component
cc$no                          #number of components

g3 <- induced_subgraph(g2, which(cc$membership == which.max(cc$csize)))    #subnetwork - giant component

#png("csg_plot3.png", width=800, height=700)
plot(g3, vertex.shape="none",edge.width=1.5, edge.curved = .5, edge.arrow.size=0.5, asp=9/16,margin=-0.15)
#dev.off()
```

### 節點的中心度
##Node centrality

```{r}
head(degree(g3, mode="in"))                     #node indegree
head(degree(g3, mode="out"))                    #node outdegree
ind <- strength(g3, mode="in")            #node indegree, using edge weights
V(g3)[order(ind, decreasing=T)[1:3]]      #top-5 nodes, based on (weighted) indegree

head(closeness(g3))                      #closeness centrality
head(betweenness(g3))                    #betweenness centrality
head(evcent(g3)$vector)                  #eigenvector centrality
```

### 凝聚力，這裡有 graph.density, transitivity (轉移關係), reciprocity(交互關係)
##Network cohesion

```{r}
graph.density(g3)           #density 
transitivity(g3)            #(global) clustering coefficient - rel. frequency connected triples close to form triangles
 
reciprocity(g3, mode="default")     #number of dyads with reciprocated (mutual) edges/number of dyads with single edge
reciprocity(g3, mode="ratio")       #total number of reciprocated edges/total number of edges
```

### 社群的辨識度
##Community identification

```{r warning=FALSE}
wt <- walktrap.community(g3)       #community identification (walktrap algorithm)
wt
length(wt)                         #number of communities
sizes(wt)                          #community sizes

#png("csg_plot4.png", width=800, height=700)
plot(wt, g3, vertex.shape="none",edge.width=1.5, edge.curved = .5, edge.arrow.size=0.5, asp=9/16,margin=-0.15)
#dev.off()
```









